---
title: "Custom Implementation of RCTD"
author: "Cody Slater"
date: "April 24, 2022"
---

```{r setup}
library(spacexr)
library(Matrix)
library(doParallel)
library(ggplot2)
```

```{r, include = TRUE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  out.width = "100%"
)
```

## Introduction

The following is a script to apply the custom gene definition from single cell datasets to spatial Visium data.

## Load data for comparison between baseline labeling and enhanced microglial labeling on LPS 1

Load all the required info for the spatial data into a puck and visualize the nUMIs distributed across the sample.

```{r Run RCTD on baseline grouping}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_LPS1.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_LPS1.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.75) 

```

```{r}
counts <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_orig.csv") # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_orig.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_orig.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_full.rds')

myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_full.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte')
print(head(norm_weights[,cell_types])) # observe weight values
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'Microglia_Macrophages'], ylimit = c(0,0.5), title ='plot of Microglia weights', size=1.5) # plot microglia/macrophage generic
```

Load all the required reference data into the reference object and run RCTD

```{r Run RCTD on stratified microglial groupings}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_LPS1.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_LPS1.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.5) 
```

```{r}
counts = as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_micro_groups.csv")) # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_micro_groups.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_micro_groups.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_LPS1_w_groups.rds')

myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_LPS1_w_groups.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte','homeostatic','intermediate activated','activated')
print(head(norm_weights[,cell_types])) # observe weight values

```

```{r}
myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_LPS1_w_groups.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte','homeostatic','intermediate activated','activated')
#(head(norm_weights[,cell_types])) # observe weight values

plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'homeostatic'], ylimit = c(0,0.5), title ='plot of homeostatic microglia weights with LPS', size=1.75) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'intermediate activated'], ylimit = c(0,0.5), title ='plot of intermediate activated microglia weights with LPS', size=1.75) # plot of intermediate activated microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'activated'], ylimit = c(0,0.5), title ='plot of activated microglia weights with LPS', size=1.75) # plot of activated microglia 
```

## Load data for comparison between baseline labeling and enhanced microglial labeling on SALINE 2 sample

Load all the required info for the spatial data into a puck and visualize the nUMIs distributed across the sample.

```{r}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_SAL2.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_SAL2.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.5) 

counts = as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_micro_groups.csv")) # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_micro_groups.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_micro_groups.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_SAL2_w_groups.rds')
```

```{r}
myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_SAL2_w_groups.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte','homeostatic','intermediate activated','activated')
#(head(norm_weights[,cell_types])) # observe weight values

plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'homeostatic'], ylimit = c(0,0.5), title ='plot of homeostatic microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'intermediate activated'], ylimit = c(0,0.5), title ='plot of intermediate activated microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'activated'], ylimit = c(0,0.5), title ='plot of activated microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
```

## Load data for comparison between baseline labeling and enhanced microglial labeling on SALINE 3 sample

Load all the required info for the spatial data into a puck and visualize the nUMIs distributed across the sample.

```{r with generic groups}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_SAL3.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_SAL3.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.5) 

counts = as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_orig.csv")) # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_orig.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_orig.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_SAL3_generic.rds')
```

```{r with generic groups}
myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_SAL3_generic.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte')
#(head(norm_weights[,cell_types])) # observe weight values

plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'Microglia_Macrophages'], ylimit = c(0,0.5), title ='plot of generic microglia/macrophage weights with saline', size=1.1) # plot of homeostatic microglia 
```

```{r}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_SAL3.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_SAL3.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.5) 

counts = as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_micro_groups.csv")) # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_micro_groups.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_micro_groups.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_SAL3_w_groups.rds')
```

```{r}
myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_SAL3_w_groups.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte','homeostatic','intermediate activated','activated')
#(head(norm_weights[,cell_types])) # observe weight values

plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'homeostatic'], ylimit = c(0,0.5), title ='plot of homeostatic microglia weights with saline', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'intermediate activated'], ylimit = c(0,0.5), title ='plot of intermediate activated microglia weights with saline', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'activated'], ylimit = c(0,0.5), title ='plot of activated microglia weights with saline', size=1.25) # plot of homeostatic microglia 
```

## Load data for comparison between baseline labeling and enhanced microglial labeling on LPS 2 sample

Load all the required info for the spatial data into a puck and visualize the nUMIs distributed across the sample.

```{r}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_LPS2.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_LPS2.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.5) 

counts = as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_micro_groups.csv")) # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_micro_groups.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_micro_groups.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_LPS2_w_groups.rds')
```

```{r}
myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_LPS2_w_groups.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte','homeostatic','intermediate activated','activated')
#(head(norm_weights[,cell_types])) # observe weight values

plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'homeostatic'], ylimit = c(0,0.5), title ='plot of homeostatic microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'intermediate activated'], ylimit = c(0,0.5), title ='plot of intermediate activated microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'activated'], ylimit = c(0,0.5), title ='plot of activated microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
```

## Load data for comparison between baseline labeling and enhanced microglial labeling on LPS 3 sample

Load all the required info for the spatial data into a puck and visualize the nUMIs distributed across the sample.

```{r with generic groups}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_LPS3.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_LPS3.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.5) 

counts = as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_orig.csv")) # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_orig.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_orig.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_LPS3_generic.rds')
```

```{r with generic groups}
myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_LPS3_generic.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte','homeostatic','intermediate activated','activated')
#(head(norm_weights[,cell_types])) # observe weight values

plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'Microglia_Macrophages'], ylimit = c(0,0.5), title ='plot of generic microglia/macrophage weights with LPS', size=1.25) # plot of homeostatic microglia 

```

```{r with custom groups}
### Load in/preprocess your data, this might vary based on your file type
counts <- as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_LPS3.csv")) # load in counts matrix
coords <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/coords_LPS3.csv")
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
rownames(coords) <- coords[,1]; coords[,1] <- NULL # Move first column to rownames
nUMI <- colSums(counts) # In this case, total counts per pixel is nUMI
puck <- SpatialRNA(coords, counts, nUMI)
barcodes <- colnames(puck@counts) # pixels to be used (a list of barcode names). 
plot_puck_continuous(puck, barcodes, puck@nUMI, ylimit = c(0,round(quantile(puck@nUMI,0.9))), title ='plot of nUMI', size = 1.5) 

counts = as.data.frame(readr::read_csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/counts_micro_groups.csv")) # load in cell types
rownames(counts) <- counts[,1]; counts[,1] <- NULL # Move first column to rownames
cell_types <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/cell_types_micro_groups.csv") # load in cell types
cell_types <- setNames(cell_types[[2]], cell_types[[1]])
cell_types <- as.factor(cell_types) # convert to factor data type
nUMI <- read.csv("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/nUMI_micro_groups.csv") # load in cell types
nUMI <- setNames(nUMI[[2]], nUMI[[1]])
reference <- Reference(counts, cell_types, nUMI)
myRCTD <- create.RCTD(puck, reference, max_cores = 32)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(myRCTD,'myRCTD_visium_LPS3_w_groups.rds')
```

```{r with custom groups}
myRCTD <- readRDS("~/GitHub/BMEN4480_FinalProject/Cody/reference_datasets/myRCTD_visium_LPS3_w_groups.rds")
barcodes <- colnames(myRCTD@spatialRNA@counts)
weights <- myRCTD@results$weights
norm_weights <- normalize_weights(weights)
cell_types <- c('Neuron.Slc17a6', 'Interneuron', 'Astrocyte','Microglia_Macrophages','Oligodendrocyte','homeostatic','intermediate activated','activated')
#(head(norm_weights[,cell_types])) # observe weight values

plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'homeostatic'], ylimit = c(0,0.5), title ='plot of homeostatic microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'intermediate activated'], ylimit = c(0,0.5), title ='plot of intermediate activated microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
plot_puck_continuous(myRCTD@spatialRNA, barcodes, norm_weights[,'activated'], ylimit = c(0,0.5), title ='plot of activated microglia weights with LPS', size=1.25) # plot of homeostatic microglia 
```
